### 2022-05-10 report/ms.tex

\documentclass{article}

\usepackage{amsmath}
\usepackage{geometry}
\usepackage{natbib}

\title{Proportions of Variants of Concern in NML Wastewater Data}
\author{Devan Becker, others}
\date{}

\begin{document}

\maketitle

\begin{abstract}
Wastewater is good for many things.
Others have done some variant detection via various methods.
Our method is principled, statistically sound, and cool.
Mutation lists are an NP-hard problem (not actually, but basically).
The method performs great, package is set up in a convenient way (TODO: set up package in a convenient way).
\end{abstract}

\section{Introduction}

Did you know there was a pandemic?
Wastewater has been shown to be useful for many analyses and pandemic response policies.

There are a few approaches to detection of SARS-CoV-2 variants\footnote{We use the convention that ``variant'' refers to a particular subtype of SARS-CoV-2 that is of interest to health authorities, such as the WHO's designated Variants of Concern (VoC) or Variants of Interest (VoI), and ``lineage'' refers to any subtype of SARS-CoV-2, whether it's designated a VOC, VOI, or neither.} in wastewater.
PCR-based search for particular variants when there are very specific mutations that are uniquely defining and easy to target.
For instance, the S-Gene Target Failure (SGTF) uniquely defines the BA.1 subvariant of B.1.1.529 (Omicron), but is not present in BA.2.
Detection of SGTF can be done without performing a full sequencing of a wastewater sample, and was useful in the detection of Omicron.

The remaining methods rely on full sequencing of the wastewater sample.
Research is still being performed on optimal sequencing strategies.
For a thorough discussion, do your own literature review. TODO: References for wastewater sequencing.
Under the ARTIC protocol (one of the most commonly used sequencing strategies), sequencing is performed using 48 amplicon regions of approximately 400bp each.
This results in a short read file of many reads that may or may not come from the same genome\footnote{It is essentially impossible to obtain and sequence any complete genomes from wastewater}.
Most common processing pipelines result in a list of mutations (relative to some reference genome, often Wuhan-1), along with the number of times those mutations were observed.
The read depth at each position relative to a reference is also obtained.
For the rest of this paper, we are primarily concerned with these data, which we will refer to as counts and coverage, respectively.

Instead of a targeted sequencing analysis (such as searching for SGTF), one can search the list of mutations for lineage-defining mutations.
Given a list of mutations that are believed to be unique (or sufficiently representative) to one lineage, the average of the observed frequencies\footnote{Frequency here is defined as the count divided by the coverage at that location.} of the mutations can be reported as an estimate of the abundance of that lineage \citep{chrystal's preprint}.
The counts and coverage can be seen as ``successes'' and ``sample size'' in a binomial model, which allows for something like a weighted average of frequencies along with a confidence interval.
This model can be found in GromStole, where the authors incorporate overdispersion for a more robust estimate of the confidence interval limits.

We are aware of two methods for ``deconvolution'' or ``de-mixing'' of mutations and variants.
Both methods are based on the idea that the frequency of a mutation should be approximately equal to the sum of the proportions of lineages present in the wastewater that contain that mutation.
For instance, if mutation A has a frequency of 0.75 and both Lineage 1 and Lineage 2 are known to have mutation A, then the porportion of Lineage 1 plus the proportion of Lineage 2 should (ideally) sum to 0.75 (but in practice there will be some error).
Given a sufficiently large list of mutations and assuming an appropriate error structure, it is possible to determine the relative frequencies of different lineages.

Freyja \citep{freyja}{part of iVar} and Alcov implement a linear regression based approach to deconvolution of mutation frequencies.
Alcov uses an unconstrained linear regression approach, which is both easy to interpret and provides reasonable estimates despite not constraining the proportions to be positive or to sum to 1.
Freya uses a robust approach and includes constraints to ensure positivity and sum-to-one.
Both methods are parts of sophisticated sequence processing pipelines.

In this paper, we introduce and implement an approach to deconvolution based on the binomial model to properly accommodate the error structure and implements positivity and sum-to-one constraints for valid estimates of proportions of lineages.
The method is implemented in both a Bayesian and a constrained optimization framework, both of which have advantages for different situations.
This method is a statistically sound and principled approach to estimation of proportions of lineages and is implemented in a user-friendly R package.

\section{Background}

\subsection{The Variant Matrix}

We define the variant matrix $M$ as a matrix where the columns represent mutations and the rows represent variants.
The $i,j$ entry is a 1 if mutation $j$ is present in variant $i$ and 0 otherwise.
This provides a very convenient data structure for analysis which allows for matrix multiplication.
We emphasize the importance of the variant matrix.
The variants that are included and the mutations associated with those variants are of paramount importance, and we stress that poor specification can have a large impact on the results of any deconvolution algorithm.

\subsection{Alcove and Freyja}

In this section, we describe the methods of Alcove and Freyja in general terms.
While not emphasizing its use, both approaches use a form of what we call the variant matrix. 
Both methods rely on the functional form:

\begin{equation}
\text{\underline{Frequency}} = \underline \rho M + \underline \epsilon \label{OLS}
\end{equation}

\noindent where \underline{Frequency} is the vector of frequencies, $\underline \rho$ is the vector of proportions of each lineage, $M$ is the variant matrix, and $\underline \epsilon$ is the vector of errors.
Alcov uses a closed form solution to ordinary least squares optimization of $\underline \rho$ in Equation \ref{OLS}.
Freyja imposes positivity and sum-to-one constraints on $\rho$ and implements a bootstrap algorithm to robustly estimate a confidence interval around the estimates.

\section{ProVoC}

We introduce ProVoC, which implements a binomial error structure that properly incorporates the information from coverage and imposes the positivity and sum-to-one constraints.
We first demonstrate the features of this method using a motivating example, then provide the specificiation of the model.

\subsection{Motivating Example}

% Use the example from the presentation. It's good.
% This might be too "teachy" for a paper, though.

\subsection{Model Specification}

Our model assumes a binomial distribution for the counts (not the frequencies), where the binomial probability is calculated according to $\underline \rho X$ and the sample size is the coverage at the relevant position on the reference:

$$
\begin{aligned}
\text{Count}_i &\sim \text{Binom}(p_i, \text{Coverage}_i)\\
\underline p &= \underline M \times \rho
\end{aligned}
$$

In this formulation, $\rho$ is a latent vector that must be estimated.
Estimation is performed in two different ways.
A Bayesian model allows for inspection of posterior distributions and posterior covariance, as well as providing a basis for more complex analyses (see Conclusions and Future Work).
Note that the Bayesian model specification includes a Dirichlet prior to impose the positivity and sum-to-one constraints on $\rho$.
A constrained optimization is also implemented, which fits much faster and a bootstrap resampling scheme allows for estimation of confidence intervals.

\subsection{Residual Analysis}

After obtaining an estimate of $\underline\rho$, which we label $\hat{\underline\rho}$, it is possible to obtain the expected counts of each mutation using the equation:
$$\widehat{\text{\underline{Count}}} = \text{\underline{Coverage}}\odot\left(M\times\hat{\underline\rho}\right)$$

\noindent where $\odot$ represents Hadamard (element-wise) multiplication.

These residuals can hold important clues for possible omitted lineages.

\section{Choosing Mutation Lists}

\section{Application}

\section{Conclusions}

\end{document}